An elaborate data set on human gait and the effect of mechanical perturbations

### 2014-12-10 paper.tex

  Here we share a rich gait data set collected from fifteen subjects walking at
  three speeds on an instrumented treadmill. Each trial consists of 120 seconds
  of normal walking and 480 seconds of walking while being longitudinally
  and/or laterally perturbed during each stance phase with pseudo-random
  fluctuations in the speed of the treadmill belt. A total of approximately 1.5
  hours of normal walking ($>5000$ gait cycles) and 6 hours of perturbed walking
  ($>20,000$ gait cycles) is included in the data set. We provide full body raw
  marker locations and ground reaction loads in addition to a presentation of
  processed data that includes gait landmarks, 2D joint angles, angular rates,
  and joint torques and the software used for the computations. The protocol is
  described in detail along with additional elaborate meta data. This data can
  likely be useful for validating or generating mathematical models that are
  capable of simulating normal periodic gait and non-periodic, perturbed gaits.

@@

\begin{itemize}
  \item A ForceLink R-Mill which has dual 6 degree of freedom force plates,
    independent belts for each foot, and lateral and pitch motion capabilities.
  \item A 10 Camera Motion Analysis motion capture system which includes Osprey
    cameras, the Cortex software, and hardware for collecting analog and camera
    data simultaneously. Delsys wireless EMG + 3D Accelerometers.
  \item Motek Medical’s D-Flow software and visual display system.
\end{itemize}

Cortex alone is capable of delivering data from the cameras along with force
plates, and analog sensors (EMG/Accelerometer) through a National Instruments
XXXX data acquisition unit. D-Flow is required to collect data from any digital
sensors and the treadmill's motion (lateral, pitch, and belts). D-Flow can
output multiple types of files which contain the different data.

@@

@@

\begin{figure*}
  \includegraphics{figures/input_vs_output.pdf}
  \caption{Treadmill belt velocity input signals (purple) and recorded output
    velocities (blue) for average belt speeds of 0.8, 1.2, and 1.6 m/s,
    respectively}
  \label{fig:input_output}
\end{figure*}

\begin{figure*}
  \includegraphics{figures/frequency_analysis.pdf}
  \caption{Frequency spectrum of the treadmill belt velocity input signal
    (purple) and the recorded output velocity (blue) for average belt speeds of
    0.8, 1.2, and 1.6 m/s, respectively}
  \label{fig:freq_analysis}
\end{figure*}

>>>>

  Here we share a rich gait data set collected from fifteen subjects walking at
  three speeds on an instrumented treadmill. Each trial consists of 120 seconds
  of normal walking and 480 seconds of walking while being longitudinally
  and/or laterally perturbed during each stance phase with pseudo-random
  fluctuations in the speed of the treadmill belt. A total of approximately 1.5
  hours of normal walking ($>5000$ gait cycles) and 6 hours of perturbed walking
  ($>20,000$ gait cycles) is included in the data set. We provide full body
  marker locations and ground reaction loads in addition to a presentation of
  processed data that includes gait landmarks, 2D joint angles, angular rates,
  and joint torques and the software used for the computations. The protocol is
  described in detail along with additional elaborate meta data. This data can
  likely be useful for validating or generating mathematical models that are
  capable of simulating normal periodic gait and non-periodic, perturbed gaits.

@@

\begin{itemize}
  \item A ForceLink R-Mill treadmill which has dual 6 degree of freedom force
    plates, independent belts for each foot, and motion capabilities (lateral,
    pitch).
  \item A 10 camera Motion Analysis motion capture system which includes Osprey
    cameras, the Cortex software (3.1.1.1290), and hardware for collecting
    analog and camera data simultaneously.
  \item National Instruments USB-6255 data aquisition unit.
  \item Four Sparkfun ADXL330 Triple Axis Accelerometer Breakout boards
    attached to the treadmill.
  \item Motek Medical’s D-Flow software (3.16.1 to 3.16.2) and visual display
    system.
\end{itemize}

Cortex alone is capable of delivering data from the cameras along with force
plates, and analog sensors (EMG/Accelerometer) through a National Instruments
USB-6255 data acquisition unit. D-Flow is required to collect data from any
digital sensors and the treadmill's motion (lateral, pitch, and belts). D-Flow
can output multiple types of files which contain the different data.

@@

\section*{Protocol}
%
The experimental protocol consisted of static measurements and walking on the
treadmill for 10 minutes under unperturbed and perturbed conditions. Before a
set of trials on the same day the following happened:
%
\begin{itemize}
  \item Calibration of the motion capture system.
  \item Subject changes into atheltic shoes, lyrca shorts, sports bra, baseball
    cap, and rock climbing harness.
  \item All 47 markers are applied directly to the skin except for the heel,
    toe, and head markers.~\footnote{The sacrum and rear pelvic markers were
    placed on the shorts for a small number of the subjects}.
  \item Subjects self-reported age, gender, and mass.
  \item Height is measured by the experimentalist.
  \item Four reference photographs (front, back, right, left) are taken of
    subjects marker locations.
\end{itemize}

After a briefing by the expermentalist on the trial protocol the subject
follows the verbal instructions of the experimentalist and the on-screen
instructions from the video display in front of the subject. The protocal goes
as so:
%
\begin{enumerate}
  \item Subject steps onto the treadmill and markers are identified with
    Cortex.
  \item The safety cable is attached loosely to the rock climbing harness such
    that no undue forces are acting on the subject during walking, but that the
    harness will prevent a full fall.
  \item The subject starts by stepping on sides of treadmill so that feet don't
    touch the force plates and the force plate signals are zero'd.
  \item Once notified by the screen, the subject stands in the initialization
    pose: standing straight up, looking forward, arms out by their sides (~45
    degrees) and the event is manually recorded by the operator.
  \item A countdown to the first normal walking phase is displayed. At the end
    of the countodwn the treadmill starts at the specified speed and the
    subject is instructed to walk normally, to focus on the road in the
    screen, and not to look at their feet.
  \item After 1 minute of normal walking, the longitudinal perturbation phase
    begins.
  \item After 8 minutes of walking under the influence of the perturbations,
    the normal walking phase begins.
  \item After 1 minute, a countdown is shown and the treadmill decelerates to a
    stop.
  \item The subject is instructed to step off of the force plates for 10
    seconds.
  \item The subject could take a break before each additional trial.
\end{enumerate}

\begin{figure*}
  \centering
  \includegraphics{figures/input_vs_output.pdf}
  \caption{Treadmill belt velocity input signals (purple) and recorded output
    velocities (blue) for average belt speeds of 0.8, 1.2, and 1.6 m/s,
    respectively}
  \label{fig:input_output}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics{figures/frequency_analysis.pdf}
  \caption{Frequency spectrum of the treadmill belt velocity input signal
    (purple) and the recorded output velocity (blue) for average belt speeds of
    0.8, 1.2, and 1.6 m/s, respectively}
  \label{fig:freq_analysis}
\end{figure*}

@@
